logistic regression

### 2021-07-01 scripts/48_16s_beta_diversity.Rmd
every movement is so hard
update
udpate

all_pheno <- bind_rows(healthy %>% 
  select(sampleid) %>% 
  mutate(grp = 'healthy', center = 'healthy'),
  stb %>% select(sampleid, center)  %>% 
    mutate(grp = 'CART') %>% 
    select(sampleid, grp, center)
  ) %>% 
  ungroup

@@

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
  labs(title = 'PCOA of healthy and CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA of healthy and CART patients.pdf')

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'center') +
  labs(title = 'PCOA of healthy and CART patients center') +
  #theme_void() +
  ggsave('../figs/PCOA of healthy and CART patients centetr.pdf')

@@

>>>>

all_pheno <- bind_rows(healthy %>% 
  select(sampleid) %>% 
  mutate(grp = 'healthy', center = 'healthy'),
  stb %>% select(sampleid, center)  %>% 
    mutate(grp = 'CART') %>% 
    select(sampleid, grp, center)
  ) %>% 
  ungroup %>% 
  inner_join(asv_alpha_diversity_ag %>% 
               select(path_pool, sampleid))

@@

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
  labs(title = 'PCOA of healthy and CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'center') +
  labs(title = 'PCOA of healthy and CART patients center') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients centetr.pdf')

@@

```{r}
# a pcoa at asv level for only M and P
cts <- get_counts_subset(c(stb$sampleid))

keepa <- cts %>% 
  filter(count_relative > 0.0001) %>% 
  count(asv_key) %>% 
  filter(n > floor(nsamp * 0.1)) %>% 
  pull(asv_key)

mp <- bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct(sampleid, .keep_all = T)  %>% 
  mutate(pool = str_extract(path_pool, 'Sample.+/'))

mp %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'pool', shape = 'center') +
  labs(title = 'PCOA of healthy and CART patients') +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 5, height = 6)
```

### 2021-07-01 scripts/57_16s_lefse.Rmd
udpate

# gather the species level taxa in the lefse significant results
res_all <- res %>% 
  set_names(res) %>% 
  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  keep(~ nrow(.) > 0) %>% 
  bind_rows(.id = 'res') %>% 
  mutate(res = str_replace(res, '^.+//',''),
         res = str_replace(res, '_asv.+$',''))  %>% 
  rename(grp = res) %>% 
  filter(grp %in% c('pull_cr_d100','pull_toxicity')) %>% 
  mutate(feature = str_replace_all(feature, '\\.','\\|')) %>% 
  #split(., list(.$grp, .$direction)) %>% 
  #map_dfr(~ top_n(x = ., n = 4, wt = score) %>% arrange(-score)) %>% 
  filter(str_detect(feature, 's__.+$')) %>% 
  filter(!str_detect(feature, 'a__.+$')) %>% 
  mutate(feature = str_replace(feature, '^.+s__','')) %>% 
  mutate(feature = str_replace(feature, '_Clostridium_', '[Clostridium]')) %>% 
  ungroup()

@@

joined <- cts_spp %>% 
  inner_join(res_all, by = c('species' = 'feature'))

@@

joined  %>%
  filter(grp == 'pull_cr_d100') %>% 
  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
  facet_wrap(direction ~ species) +
  ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)

@@

```{r}
joined  %>%
  filter(grp == 'pull_toxicity') %>% 
  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
  facet_wrap(direction ~ species) +
  ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)
```

>>>>

# gather the species level taxa in the lefse significant results
res_all <- res %>% 
  set_names(res) %>% 
  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  keep(~ nrow(.) > 0) %>% 
  bind_rows(.id = 'res') %>% 
  mutate(res = str_replace(res, '^.+//',''),
         res = str_replace(res, '_asv.+$',''))  %>% 
  rename(grp = res) %>% 
  filter(grp %in% c('pull_cr_d100','pull_toxicity')) %>% 
  mutate(feature = str_replace_all(feature, '\\.','\\|')) %>% 
  #split(., list(.$grp, .$direction)) %>% 
  #map_dfr(~ top_n(x = ., n = 4, wt = score) %>% arrange(-score)) %>% 
  # filter(str_detect(feature, 's__.+$')) %>% 
  # filter(!str_detect(feature, 'a__.+$')) %>% 
  filter(str_detect(feature, 'g__.+$')) %>% 
  filter(!str_detect(feature, 's__.+$')) %>% 
  mutate(feature = str_replace(feature, '^.+g__','')) %>% 
  mutate(feature = str_replace(feature, '_Clostridium_', '[Clostridium]')) %>% 
  ungroup()

@@

cts_genus <- cts_ %>% 
  full_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key)  %>% 
  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(Relab = sum(relab)) %>% 
  select(sampleid, genus, Relab) %>% 
  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid') %>% 
  ungroup()

joined <- cts_genus %>% 
  inner_join(res_all, by = c('genus' = 'feature'))

@@

joined  %>%
  filter(grp == 'pull_cr_d100') %>% 
  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
  facet_wrap(direction ~ genus) +
  ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)

@@

```{r}
joined  %>%
  filter(grp == 'pull_toxicity') %>% 
  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
  facet_wrap(direction ~ genus) +
  ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)
```

### 2021-07-01 scripts/62_logistic_regression_version.Rmd
udpate

gtox <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gtox_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

@@

gCR <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

>>>>

gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) 
```

@@

gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

# log 10 transformed version of the genera relab

```{r}
gmeta_log10 <- meta  %>% 
  inner_join(genera) %>% 
  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)

```

```{r}
g_tox <- glm(
  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta_log10, 
  family = binomial(link = "logit")
)
summary(g_tox)
gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
gtox_log <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gtox_log10_5.pdf'), width = 3, height = 8, dpi = 300)
```

```{r}
g_CR <- glm(
  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta_log10, 
  family = binomial(link = "logit")
)
summary(g_CR)
gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
gCR_log <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) 
```

# side-by-side comparison of the CLR and log10 transformed

```{r}
tox_all <- ggarrange(gtox_log, gtox_clr, nrow = 1) +
    ggsave(str_glue('../figs/amplicon/gtox_all_5.pdf'), width = 6, height = 8, dpi = 300)

cr_all <- ggarrange(gCR_log, gCR_clr, nrow = 1) +
    ggsave(str_glue('../figs/amplicon/gCR_all_5.pdf'), width = 6, height = 8, dpi = 300)
```
